\begin{abstract}
As Generative AI grows in popularity and usage day by
day, the effortless creation of hyper-realistic manipulated
media, commonly referred to as 'deepfakes', presents a challenge to verifying digital authenticity today. Historically, researchers have relied on training massive neural networks
to spot spatial and semantic inconsistencies directly within
raw images as their training set. Although this approach
worked well before for the early deepfakes, these visual
clues are quickly disappearing as generative models are
learning to blend images more flawlessly. In addition, relying on such massive networks creates an enormous computational burden to train on the raw RGB images.

In this study, however, I explored a fundamentally different approach to digital forensics. Instead of just analyzing the visual content looking for clues in the image, I chose to investigate the digital fingerprint left behind by the way the image was saved. To do this, I used Error Level Analysis (or ELA) to reveal hidden compression artifacts and 'residues'. When an image is manipulated or spliced, the altered regions usually have a different JPEG compression history than the untouched background. ELA highlights these subtle differences, displaying them as distinct, high-frequency visual noise.

I applied this technique to a 10,000-frame subset of
the FaceForensics++ dataset by converting standard images into detailed ELA heatmaps. Recognizing the shift in domain between natural images and noise-based heatmaps,
I fine-tuned an EfficientNet-B0 architecture using a strategy of varying learning rates. I completely unfroze the pre-trained base, which allowed the lower-level convolutional filters to specifically adapt to the compression grids. As a result, my optimized model achieved a strong validation accuracy of 86.40\% alongside an F1-score of 86.11\%. Ultimately, my findings demonstrate that compression artifact patterns can serve as highly reliable and lightweight features for detecting deepfakes.
\end{abstract}
