\section{Experiments and Results}
\label{sec:experiments}

To rigorously evaluate my Error Level Analysis and EfficientNet-B0 pipeline, I established a strict experimental protocol. I partitioned my 10,000-image dataset into an 80/20 split. This gave me 8,000 images exclusively for training and reserved 2,000 unseen images for validation. To match the input expectations of the EfficientNet architecture, I uniformly resized all images to $224\times224$ pixels.

\subsection{Training Setup and Hyperparameters}

I trained my models on a local workstation equipped with an NVIDIA RTX 3060 Ti GPU, leveraging CUDA for hardware acceleration. Given the relatively small size of my dataset, I was extremely conscious of the risk of spatial overfitting. It is very easy for a model to simply memorize the exact position of watermarks or static noise patterns in the training data.

To combat this, I aggressively employed data augmentation. During training, images were subjected to random horizontal flips and random rotations of up to 15 degrees. This forced the network to genuinely learn the fundamental texture of compression anomalies rather than relying on where those anomalies appeared on the screen.

The network was optimized using the Adam optimizer. To ensure smooth convergence and prevent the network from overshooting local minima during the final stages of training, I wrapped the optimizer in a Cosine Annealing learning rate scheduler. I ran my training loops for a maximum of 10 epochs, monitoring the validation loss at the end of each epoch so I could save the absolute best-performing model weights.

\subsection{Baseline Evaluation: The Failure of Frozen Weights}

As I alluded to in my methodology, my first set of experiments served as a baseline to highlight just how difficult the ELA domain shift truly is. I trained a model where the EfficientNet-B0 feature extractor was completely frozen.

The results of this baseline run were incredibly poor. The model struggled to converge, and testing on the validation set yielded an F1-score of roughly 58.37\%. Given that my dataset was perfectly balanced, a 58\% score is only marginally better than a random coin toss. The frozen ImageNet filters simply could not extract meaningful patterns from the high-frequency ELA noise grids. This baseline definitively proved that standard transfer learning protocols fall short when migrating from natural RGB imagery to forensic, noise-based representations.

\subsection{Final Results: Full Fine-Tuning}

Following the failure of the frozen baseline, I executed my primary experimental strategy: full network fine-tuning utilizing my proposed differential learning rates. The impact of unfreezing the lower-level convolutional layers was both immediate and dramatic. The model exhibited rapid convergence, demonstrating a steep decline in training loss within just the first three epochs.

By the ninth epoch, the network hit its peak performance on the validation set. The final, optimized model weights were saved, and the corresponding metrics are detailed in Table~\ref{tab:results}.

\begin{table}[h]
\centering
\caption{Final Validation Metrics}
\label{tab:results}
\begin{tabular}{@{}lc@{}}
\toprule
\textbf{Metric} & \textbf{Score (\%)} \\ \midrule
Accuracy & 86.40 \\
Precision & 88.00 \\
Recall & 84.30 \\
F1-Score & 86.11 \\ \bottomrule
\end{tabular}
\end{table}

These metrics tell a compelling story about the efficacy of my pipeline. Achieving an overall accuracy of 86.40\% is highly significant. It confirms that compression artifacts alone, stripped entirely of their semantic visual context, serve as highly discriminatory features for exposing digital forgery.

Furthermore, I carefully analyzed the precision and recall scores. In the context of deepfake detection, a high false-positive rate is severely detrimental; continually flagging authentic media as fake quickly erodes user trust in the detection system. My model achieved a notably high precision score of 88.00\%. This indicates a strong resistance to false positives, meaning that when my system flags an image as a deepfake, there is a very high probability that it has genuinely been manipulated. The slightly lower recall score of 84.30\% suggests that the model occasionally misses highly sophisticated fakes where the generator's compression matching is exceptionally good. However, the overall F1-score of 86.11\% demonstrates an excellent balance of reliability and sensitivity.
