\section{Conclusion and Future Work}
\label{sec:conclusion}

In this paper, we presented a highly efficient and targeted approach to deepfake detection. By stepping away from the computationally expensive trend of massive RGB semantic analysis, we demonstrated that looking at the digital physics of an image is an incredibly viable alternative. 

Our pipeline, which isolates JPEG compression inconsistencies via Error Level Analysis, successfully forces neural networks to act as digital forensic analysts. However, getting a modern CNN to understand this abstract data was not trivial. Our early failures highlighted the stark domain shift between natural imagery and compression grids. Ultimately, overcoming this shift required full network fine-tuning and the careful application of differential learning rates, ensuring our EfficientNet-B0 architecture retained its fundamental edge-detection capabilities while adapting to high-frequency noise. The end result---a lightweight model capable of differentiating authentic from manipulated media with an 86.40\% accuracy rate on the FaceForensics++ dataset---validates our core hypothesis. 

Looking forward, there are several exciting avenues for future research. While our model performs exceptionally well by looking solely at compression artifacts, it completely ignores the semantic cues that still trip up less sophisticated generative models. We envision developing a dual-branch architecture that processes both the raw RGB image and the ELA heatmap simultaneously. By fusing the spatial-semantic features from the RGB branch with the digital-forensic features from the ELA branch, we could theoretically build a hybrid detector capable of identifying deepfakes across a much wider spectrum of generative quality.
